---
title: "R Notebook"
output: html_notebook
author : "Dixitha Kasturi"
---


```{r}

library(dplyr)
library(Hmisc)
library(tidyverse)
library(cluster)
library(factoextra)
library(gridExtra)

```
__Loading Data__

```{r}
setwd("C:/Users/kastu/Desktop/Syracuse/Spring22/IST707-AML/Week4/Assignment")
papers <- read_csv("HW4-data-fedPapers85.csv")


```


__Data Exploration__

```{r}
str(papers)

```

```{r}

dim(papers)

```

* 85 papers with 72 attributes. Out of which the first two columns have the author name and the 2nd one has the actual file name of the paper. The other columns are counts of characters/words accoridng to the name of the column.

* Name of author and filename are both nominal attributes. This can be seen through the describe function.

```{r}
describe(papers[,0:2])
```

```{r}
sapply(papers, function(x) sum(is.na(x)))
```


* We see that the data is clean and has no missing values. There are 3 different authors and few combined by Hamilton and Madison and 85 text files. Out of these 85 we have 11 disputed records. We are supposed to find which cluster tehse files would belong to.
* 51 essays written by Hamilton
* 15 by Madison
* 3 by Hamilton and Madison
* 5 by Jay


* Clustering Analysis
* We will do 2 types of clustering. Kmeans and Hierarcheal clustering


Method 1 : Kmeans
* We want 3 clusters as we have 3 different authors.
* We will use the columns which have counts of words to understand the document distribution, as authors have a style of writing, we subset the data
* We then scale our data so that clustering doesnt depend only on one attribute.


```{r tidy = True}
#papers_sub <- select(papers, col = -c("author","filename"))
papers_sub <- papers[,3:72]
papers_sub <- scale(papers_sub)

set.seed(1000)
km3 <- kmeans(papers_sub,centers = 3)

#km4 <- kmeans(papers_sub,4)
km3$size
#km4$size
clusplot(papers_sub, km3$cluster, color = T, shade = F, labels = 0, lines = 0)
```


* We see that the sizes of the clusters dont match to what we expected,even if we conider HM as a different cluster. We plot the clusters to see how clusters are exactly formed

```{r}
set.seed(20)
fviz_cluster(km3, data = papers_sub, pointsize = 1, labelsize = 8.5,show.clust.cent = TRUE, ggtheme = theme_bw(), main = "3 clusters")
```

```{r}
#fviz_cluster(km4, data = papers_sub, pointsize = 1, labelsize = 8.5,show.clust.cent = TRUE, ggtheme = theme_bw(), main = "4 clusters")

```


* We clearly see how the clusters overlap and the documents on manual verification are not clustered the way they are supposed to. Even if we change the number of clusters, it still didnt give good results.

* We choose the optimum number of clusters that need to be formed using Elbow method

```{r}

# Elbow method

set.seed(20)
fviz_nbclust(papers_sub, kmeans, method = "wss", k.max=5)

```



* It is still not very clear, as to how many clusters should be taken. Hence we stick with 3.

```{r}
clusters <- cbind(papers,km3$cluster)

final <- clusters %>% group_by(author,km3$cluster) %>% summarise(count =  n())
final

```


Method 2:

Hierarceal clustering
a) using complete method
b) using average method
c) Using single method

```{r}

h_complete <- hclust(dist(papers[, 3:72]), method = 'complete')
h_complete
plot(h_complete, hang = -1, cex = 0.6, label = papers$author,
      main = "Papers Cluster using Complete")


```

```{r}

h_avg <- hclust(dist(papers[, 3:72]), method = 'average')
h_avg
plot(h_avg, hang = -1, cex = 0.6, label = papers$author,
     main = "Papers Cluster using Average")

```


```{r}
h_single <- hclust(dist(papers[, 3:72]), method = 'single')
h_single
plot(h_single, hang = -1, cex = 0.6, label = papers$author,
     main = "Papers Cluster using Single method")

```

* When we use complete and average methods in hierarcheal clustering, we see that Jay was grouped together but in single method the results were a little similar to what we saw from kmeans. While it is unclear, going by the majority, we see that HAC performed better in clustering the files. Jay were grouped. In complete and average method - there was overlap between hamilton, madison and disputed articles. So we can be sure here that the articles were written by either of them. The same went for single method, but the only difference was that jay articles were in different clusters.


```{r}
required_authors= papers$author %in% c("Hamilton", "Madison", "HM", "dispt")
papers_sample <- papers[required_authors,]
papers_2 <- papers[required_authors,3:72]
papers_2 <- scale(papers_2)
#kmeans :

set.seed(1000)
km3_2 <- kmeans(papers_2,centers = 2, nstart = 25, iter.max = 10)

km3_2$size

clusplot(papers_2, km3_2$cluster, color = T, shade = F, labels = 0, lines = 0)

set.seed(20)
fviz_cluster(km3_2, data = papers_2, pointsize = 1, labelsize = 7,
             show.clust.cent = TRUE, ggtheme = theme_bw(), main = "3 clusters")


```

```{r}
clusters_2 <- cbind(papers_sample,km3_2$cluster)

final_2 <- clusters_2 %>% group_by(author,km3_2$cluster) %>% summarise(count =  n())
final_2

```

```{r}

sort(abs(km3_2$centers[1,]-km3_2$centers[2,]))

```


The blue cluster is mostly hamilton and the red is mostly madison.
The joint authorship papers(64,65,63) are located in the red cluster far from the center of the red cluster.

Method 2 : Using hierarchical clustering:

```{r}
h_complete2 <- hclust(dist(papers_sample[, 3:72]), method = 'complete')
h_complete2
plot(h_complete1, hang = -1, cex = 0.6, label = papers_sample$author,
      main = "Papers Cluster using Complete")

```


```{r}
h_avg2 <- hclust(dist(papers_sample[, 3:72]), method = 'average')
h_avg2
plot(h_avg2, hang = -1, cex = 0.6, label = papers_sample$author,
     main = "Papers Cluster using Average")

```


```{r}
h_single2 <- hclust(dist(papers_sample[, 3:72]), method = 'single')
h_single2
plot(h_single2, hang = -1, cex = 0.6, label = papers_sample$author,
     main = "Papers Cluster using Single method")


```


According to this report, the disputed papers belong to Madison from both HAC( all 3 methods) & Kmeans analysis.
We see that out of all the attributes the word “upon” is the most useful for clustering and differentiating
the papers. The centroid values for the ‘upon’ dimension are the farthest.

If we included the papers from Jay, there was a lot of distortion and the results were not accurate. When these papers were excluded exactly according to the claims of the report,we see that the disputed files leaned more towards madison. So we can say that the articles belonged to madison.


